---
title: "Analyses for Susa (2016) mark-recapture study"
author: "Johanna Kiss et al."
date: "2023-12-07"
output:
  html_document:
    toc: yes
    toc_depth: 4
    theme: united
  pdf_document: 
    toc: yes
    toc_depth: 4
    fig_caption: yes
    keep_tex: yes
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 18px;
  color: Gunmetal;
}
h1.title {
  font-size: 32px;
  color: Gunmetal;
  font-weight: bold;
}
h1 { /* Header 1 */
  font-size: 30px;
  color: BlueMossGreen;
}
h2 { /* Header 2 */
    font-size: 26px;
  color: DarkSlateGray;
  font-weight: bold;
}
h3 { /* Header 3 */
  font-size: 20px;
  color: DeepTeal;
  font-weight: bold;
}
h4 {/* Header 4 */
  font-size:16px;
  color: MediumTeal;
  font-weight: bold;
}
code.r{ /* Code block */
    font-size: 10px;
    color: CadetBlue;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 10px;
}
</style>



### Describing movement of different sized males and females in Lethrus beetles on field by using one model.




 Using 1 model for each response variable (e.g., number of recaptures, travelled distances...) instead of 2.

 To do that we reanalyse data and test the effect of the created 4-level factor ( female, small male, middle male, large male)


```{r}
all.dist<-read.csv(file="rel.full.data.csv",header=TRUE,nrow=FALSE)
all.capt<-read.csv(file="rel.aa.df.data.csv",header=TRUE,nrow=FALSE)
```

```{r}
head(all.dist)
head(all.capt)
```


 Previously,  based on a segmented regression on body size we determined the breakpoints, now we use them to create male categories:

```{r}
library(dplyr)
new.all.capt<-all.capt%>%
  mutate(sex.cat= case_when(sex=="m" & B.width <= 11.10 ~ 'small.male',
                            sex=="m" & B.width > 12.37 ~ 'large.male',
                            sex=="m" & B.width > 11.10 & B.width <= 12.37 ~ 'middle.male',
                            sex=="f" ~ 'female'))

```

```{r}

library(dplyr)
new.all.dist<-all.dist%>%
  mutate(sex.cat= case_when(sex=="m" & body <= 11.10 ~ 'small.male',
                            sex=="m" & body > 12.37 ~ 'large.male',
                            sex=="m" & body > 11.10 & body <= 12.37 ~ 'middle.male',
                            sex=="f" ~ 'female'))

```

 Load some packages:

```{r}
library(plotrix)
library(emmeans)
library(plyr)
library(dplyr)
library(ggplot2)
library(scales)
library(MASS)
library(lme4)
library(pbkrtest)
library(pscl)
library(lattice)
library(partR2)
library(rsq)
library(segmented)
library(interactions)
library(jtools)
library(bbmle)
library(cowplot)
```


#### New analyses on the number of recaptures

```{r}
str(new.all.capt)
```

```{r}
new.all.capt.m<-new.all.capt[new.all.capt$sex=="m",]

new.all.capt.m<-new.all.capt.m[new.all.capt.m$Tusk>0,]

str(new.all.capt.m)


custom_pch <- c(3,15,17)

plot(new.all.capt.m$Tusk~new.all.capt.m$B.width, type="n", xlab="Body width [mm]", ylab="Tusk length [mm]",las=1,  cex.axis=1.3, cex=1, cex.lab=1.7, las=1, font.lab=2.3, font.main=4.3, ylim=c(0, 5),xlim=c(9,14))
points(new.all.capt.m$Tusk~new.all.capt.m$B.width, subset=new.all.capt.m$sex.cat=="small.male", pch=3,
       col=1)
points(new.all.capt.m$Tusk~new.all.capt.m$B.width, subset=new.all.capt.m$sex.cat=="middle.male", pch=15,
       col="grey")

points(new.all.capt.m$Tusk~new.all.capt.m$B.width, subset=new.all.capt.m$sex.cat=="large.male", pch=17,
       col="darkgrey")

abline(v=c(11.10, 12.37), col=c("darkgrey", "black"), lty=2, lwd=2)



```

 Lets see the models:

```{r}
new.p.rec<-glm(no.recaptures~sex.cat,new.all.capt,family=poisson)
drop1(new.p.rec,test="Chisq")
```

```{r}
new.qp.rec<-glm(no.recaptures~sex.cat,new.all.capt,family=quasipoisson)
drop1(new.qp.rec,test="Chisq")
```

```{r}
new.zp.rec<-zeroinfl(no.recaptures~sex.cat,new.all.capt,dist='poisson')
drop1(new.zp.rec,test="Chisq")
```

```{r}
new.znb.rec<-zeroinfl(no.recaptures~sex.cat,new.all.capt,dist='negbin')
drop1(new.znb.rec,test="Chisq")
```
```{r}
new.hnb.rec<-hurdle(no.recaptures~sex.cat,new.all.capt,dist='negbin')
drop1(new.hnb.rec,test="Chisq")
```

```{r}
new.nb.rec<-glm.nb(no.recaptures~sex.cat,new.all.capt)
drop1(new.nb.rec,test="Chisq")
```

 Lets compare the models

```{r}

logLik(new.p.rec)
logLik(new.nb.rec)
logLik(new.hnb.rec)
logLik(new.znb.rec)
logLik(new.zp.rec)
logLik(new.qp.rec)
```
```{r}

AIC(new.p.rec)
AIC(new.nb.rec)
AIC(new.hnb.rec)
AIC(new.znb.rec)
AIC(new.zp.rec)
AIC(new.qp.rec)
```
 HNB and NB models seemed to be the best.

```{r}
drop1(new.hnb.rec,test="Chisq")
drop1(new.nb.rec,test="Chisq")
```

```{r}
ph.test.rec<-emmeans(new.nb.rec, ~sex.cat,type="response")
test(ph.test.rec)
pairs(ph.test.rec)

plot(ph.test.rec)


ph.test.rec2<-emmeans(new.nb.rec, ~sex.cat,type="link")
test(ph.test.rec2)
pairs(ph.test.rec2)

plot(ph.test.rec)



```

```{r}

boxplot(new.all.capt$no.recaptures~new.all.capt$sex.cat,xlab="Sex",ylab="Number of (re)captures",las=1, lwd=1,font.lab=2, cex.lab=1.3,cex.axis=1.3,font.main=2)

new1<-ggplot(new.all.capt,aes(sex.cat,no.recaptures),line.thickness=1.2) +theme_classic(base_size = 18) + geom_jitter(width=0.2,alpha=0.5)+scale_y_continuous(name="Number of (re)captures",breaks= pretty_breaks()) +theme (axis.text.x = element_text(face = "bold", size=12,colour="black"), axis.text.y = element_text(face = "bold", size = 12,colour="black"),panel.border= element_rect(color = "black",fill=NA,size=1), legend.text=element_text(size=14),legend.title=element_text(size=16))

```

```{r fig.height = 10, fig.width = 10, fig.align = "center"}

new.plot1<-new.all.capt[,c(7,12)]
new.plot1.sum<-aggregate(.~sex.cat,mean,data=new.plot1)



new1+ theme(plot.title.position = "plot")+geom_crossbar(data=new.plot1.sum, aes(ymin=no.recaptures,ymax=no.recaptures),size=0.8,width=0.4)

```

 Lets see now the effect size:

```{r}
rsq.partial(new.nb.rec,type='kl')
```

 Females and male morphs showed a small effect size: 0.014.


#### New analyses on the number of days


 Lets see the models:

```{r}
new.p.days<-glm(days~(sex.cat*no.recaptures)^2,new.all.capt,family=poisson,subset=new.all.capt$no.recaptures>0)
drop1(new.p.days,test="Chisq")

new.p.days1<-update(new.p.days,.~.-sex.cat:no.recaptures)
drop1(new.p.days1,test="Chisq")
```

```{r}
new.qp.days<-glm(days~(sex.cat*no.recaptures)^2,new.all.capt,family=quasipoisson,subset=new.all.capt$no.recaptures>0)
drop1(new.qp.days,test="Chisq")

new.qp.days1<-update(new.qp.days,.~.-sex.cat:no.recaptures)
drop1(new.qp.days1,test="Chisq")
```

```{r}
new.zp.days<-zeroinfl(days~(sex.cat*no.recaptures)^2,new.all.capt,dist='poisson',subset=new.all.capt$no.recaptures>0)
drop1(new.zp.days,test="Chisq")

new.zp.days1<-update(new.zp.days,.~.-sex.cat:no.recaptures)
drop1(new.zp.days1,test="Chisq")

```

```{r}
new.znb.days<-zeroinfl(days~(sex.cat*no.recaptures)^2,new.all.capt,dist='negbin',subset=new.all.capt$no.recaptures>0)
drop1(new.znb.days,test="Chisq")

new.znb.days1<-update(new.znb.days,.~.-sex.cat:no.recaptures)
drop1(new.znb.days1,test="Chisq")
```
```{r}
new.hnb.days<-hurdle(days~(sex.cat*no.recaptures)^2,new.all.capt,dist='negbin',subset=new.all.capt$no.recaptures>0)
drop1(new.hnb.days,test="Chisq")

new.hnb.days1<-update(new.hnb.days,.~.-sex.cat:no.recaptures)
drop1(new.hnb.days1,test="Chisq")
```

```{r}
new.nb.days<-glm.nb(days~(sex.cat*no.recaptures)^2,new.all.capt,subset=new.all.capt$no.recaptures>0)
drop1(new.nb.days,test="Chisq")

new.nb.days1<-update(new.nb.days,.~.-sex.cat:no.recaptures)
drop1(new.nb.days1,test="Chisq")
```

 Lets compare the models

```{r}

logLik(new.p.days1)
logLik(new.nb.days1)
logLik(new.hnb.days1)
logLik(new.znb.days1)
logLik(new.zp.days1)
logLik(new.qp.days1)
```

```{r}

AIC(new.p.days1)
AIC(new.nb.days1)
AIC(new.hnb.days1)
AIC(new.znb.days1)
AIC(new.zp.days1)
AIC(new.qp.days1)
```
 Here also HNB and NB models seemed to be the best.

```{r}
drop1(new.hnb.days1,test="Chisq")
drop1(new.nb.days1,test="Chisq")
```
 Its seems that sex and male morphs had no significant effect on the number of days individuals were seen on field.


```{r}

new.all.capt2<-new.all.capt[new.all.capt$no.recaptures>0,]


boxplot(new.all.capt2$days~new.all.capt2$sex.cat,xlab="Sex",ylab="Number of days between the first and last encounter",las=1, lwd=1,font.lab=2, cex.lab=1.3,cex.axis=1.3,font.main=2)
```


```{r}

new2<-ggplot(new.all.capt2,aes(sex.cat,days),line.thickness=1.2) +theme_classic(base_size = 18) + geom_jitter(width=0.2,alpha=0.5)+scale_y_continuous(name="Number of days between the first and last encounter",breaks= pretty_breaks()) +theme (axis.text.x = element_text(face = "bold", size=12,colour="black"), axis.text.y = element_text(face = "bold", size = 12,colour="black"),panel.border= element_rect(color = "black",fill=NA,size=1), legend.text=element_text(size=14),legend.title=element_text(size=16))

```

```{r fig.height = 8, fig.width = 8, fig.align = "center"}


new.plot2<-new.all.capt2[,c(5,12)]
new.plot2.sum<-aggregate(.~sex.cat,mean,data=new.plot2)



new2+ theme(plot.title.position = "plot")+geom_crossbar(data=new.plot2.sum, aes(ymin=days,ymax=days),size=0.8,width=0.4)

```

 Lets see now the effect size:

```{r}
rsq.partial(new.nb.days1,type='kl')
```


 Females and male morphs showed also a small effect size: 0.013.

```{r}
ph.test.days<-emmeans(new.nb.days1, ~sex.cat,type="response")
test(ph.test.days)
pairs(ph.test.days)
contrast(ph.test.days)

plot(ph.test.days)


ph.test.days2<-emmeans(new.nb.days1, ~sex.cat,type="link")
test(ph.test.days2)
pairs(ph.test.days2)
contrast(ph.test.days2)

plot(ph.test.days2)
```


#### New analyses of distances travelled between two encounters


 Response variables: female-male categories (4-level factor), number of days since start of the study, number of days between two encounters, and id as a random factor.

```{r}
colnames(new.all.dist)
str(new.all.dist)
```
```{r}
library(lme4)
dist.new<-lmer(log(Distance+1) ~ (sex.cat+day+diff.days)^2+(1|id),data=new.all.dist)
print(summary(dist.new),correlation=TRUE)
drop1(dist.new,test="Chisq")
```


```{r}
dist.new2<-update(dist.new,.~.-sex.cat:diff.days)
summary(dist.new2)
drop1(dist.new2,test="Chisq")
```


```{r}
dist.new3<-update(dist.new2,.~.-day:diff.days)
summary(dist.new3)
drop1(dist.new3,test="Chisq")
```

Compare "stats drop1" to "glmmTMB drop1":

```{r}
library(glmmTMB)
dist.new.tmb<-glmmTMB(log(Distance+1) ~ sex.cat:day+diff.days+(1|id),family="gaussian",data=new.all.dist)
summary(dist.new.tmb)
drop1(dist.new.tmb,test="Chisq")
```

```{r}
ph.dist.new<-emmeans::emtrends(dist.new3,~1,"diff.days")
test(ph.dist.new)
```

```{r}
effect_plot(dist.new3, pred = diff.days, interval = FALSE, plot.points = TRUE,x.label="Number of days between encounters")+ theme_apa()
```

```{r}
ph.dist.new2<-emtrends(dist.new3,~sex.cat,"day")
test(ph.dist.new2)

pairs(ph.dist.new2)

emmip(dist.new3,sex.cat~day,cov.reduce=range)
```

 Small males seem to travel and move similarly to the females.

 Lets visualise the results.

```{r fig.height = 8, fig.width = 14, fig.align = "center"}


new.dist1<-interact_plot(dist.new3, pred = day, modx = sex.cat,x.label="Number of days since the start of the study",data=new.all.dist,y.label="Log(Distance+1)",lwd=2,line.thickness=1.2,colors=c("black","grey35","grey55","grey6"),plot.points=TRUE,point.shape=T,point.size=3,interval = TRUE,  int.type = "confidence", int.width = .95,modx.labels=c("Female","Large male","Medium male","Small male")) + theme_apa()+theme_classic(base_size = 18)+theme(legend.title=element_blank(),legend.key.size = unit(3, "lines"),legend.position="right",legend.justification = "top")+guides(fill = guide_legend(override.aes = list(size=4)))+ scale_x_continuous(limits=c(0.04,71),expand=c(0,0),breaks=c(10,20,30,40,50,60)) + scale_y_continuous(limits=c(0.04,6.4), expand=c(0,0),breaks=c(1,2,3,4,5,6))



new.dist1 + theme (axis.text.x = element_text(face = "bold", size=12),  axis.text.y = element_text(face = "bold", size = 12),panel.border = element_rect(colour = "black", fill=NA, size=0.2))
```

 Create colourfull plots for graphical abstract:

```{r fig.height = 8, fig.width = 14, fig.align = "center"}


new.dist.col<-interact_plot(dist.new3, pred = day, modx = sex.cat,x.label="Number of days since the start of the study",data=new.all.dist,y.label="Log(Distance+1)",lwd=2,line.thickness=1.2,colors=c("green","brown","red","orange"),interval = TRUE, int.type = "confidence", int.width = .95) + theme_apa()+theme_classic(base_size = 18)+ scale_x_continuous(limits=c(0.04,71),expand=c(0,0),breaks=c(10,20,30,40,50,60)) + scale_y_continuous(limits=c(0.04,6.4), expand=c(0,0),breaks=c(1,2,3,4,5,6))+theme(legend.position = "none")



new.dist.col + theme (axis.text.x = element_text(face = "bold", size=16),  axis.text.y = element_text(face = "bold", size = 16),panel.border = element_rect(colour = "brown", fill=NA, size=0.5))




new.dist.col2<-interact_plot(dist.new3, pred = day, modx = sex.cat,x.label="Number of days since the start of the study",data=new.all.dist,y.label="Log(Distance+1)",lwd=2,line.thickness=1.2,colors=c("#4f5157","#fd8f24","#a81816","#89a954"),plot.points=TRUE,point.shape=T,point.size=4,interval = TRUE,  int.type = "confidence", int.width = .95,modx.labels=c("Female","Large male","Medium male","Small male")) + theme_apa()+theme_classic(base_size = 18)+theme(legend.title=element_blank(),legend.key.size = unit(3, "lines"),legend.position="right",legend.justification = "top")+guides(fill = guide_legend(override.aes = list(size=4)))+ scale_x_continuous(limits=c(0.04,71),expand=c(0,0),breaks=c(10,20,30,40,50,60)) + scale_y_continuous(limits=c(0.04,6.4), expand=c(0,0),breaks=c(1,2,3,4,5,6))



new.dist.col2 + theme (axis.text.x = element_text(face = "bold", size=12),  axis.text.y = element_text(face = "bold", size = 12),panel.border = element_rect(colour = "black", fill=NA, size=0.2))

ggsave("seasonal.movement.Figure.1.tiff", units="px",height= 3000, width= 4000,dpi=300)

```


 Lets see effect sizes:

```{r}
partR2(dist.new3, partvars= c("diff.days","sex.cat:day"),R2_type="conditional",nboot = 100, CI = 0.95)

```

```{r}
partR2(dist.new3, partvars= c("diff.days","sex.cat:day"),R2_type="marginal",nboot = 100, CI = 0.95)
```

 Semi-partial R2 of sex-size category and number of days since start of the study: marginalR2= 0.035, conditionalR2=0.155, i.e., representing medium-leveled effect.

 Lets create a plot for the first two models:

```{r plot-fig-1-2}
d.plot1<-new.all.capt[,c(7,12)]
d.plot1.sum<-aggregate(.~sex.cat,mean,data=d.plot1)

p1new<-ggplot(d.plot1,aes(sex.cat,no.recaptures),line.thickness=1.2) +theme_classic(base_size = 18)+geom_point(position=position_jitter(width=0.2),size=3,aes(shape=sex.cat,x = sex.cat, y = no.recaptures,color=sex.cat))+
scale_y_continuous(name="Number of encounters",limits=c(0,11),breaks= round(seq(min(new.all.capt$no.recaptures),max(new.all.capt$no.recaptures),by=1),1)) +  labs(x="Sex")+scale_x_discrete(labels=c("Female","Large male","Medium male", "Small male"))+theme (axis.text.x = element_text(face = "bold", size=16,colour="black"), axis.text.y = element_text(face = "bold", size = 12),panel.border= element_rect(color ="black",fill=NA,size=1))+geom_crossbar(data=d.plot1.sum, aes(ymin=no.recaptures,ymax=no.recaptures),size=0.8,width=0.6)+theme(legend.position = "none")+ scale_color_manual(values = c("black","grey35","grey55","grey6"))

p1t<-p1new+labs(title = "a)") + theme(plot.title.position = "plot")+geom_crossbar(data=d.plot1.sum, aes(ymin=no.recaptures,ymax=no.recaptures),size=0.8,width=0.4)

d.plot<-new.all.capt2[,c(5,12)]
d.plot.sum<-aggregate(.~sex.cat,mean,data=d.plot)

p2<-ggplot(d.plot,aes(sex.cat,days),line.thickness=1.2) +theme_classic(base_size = 18)+geom_point(position=position_jitter(width=0.2),size=3,aes(shape=sex.cat,x = sex.cat, y = days, color = sex.cat))+
scale_y_continuous(name="Number of days",limits=c(0,57),breaks= round(seq(min(new.all.capt2$days),max(new.all.capt2$days),by=3),1)) +  labs(x="Sex")+scale_x_discrete(labels=c("Female","Large male","Medium male", "Small male"))+theme (axis.text.x = element_text(face = "bold", size=16,colour="black"), axis.text.y = element_text(face = "bold", size = 12),panel.border= element_rect(color ="black",fill=NA,size=1))+geom_crossbar(data=d.plot.sum, aes(ymin=days,ymax=days),size=0.8,width=0.6)+theme(legend.position = "none")+ scale_color_manual(values = c("black","grey35","grey55","grey6"))


p2t<-p2+labs(title = "b)") + theme(plot.title.position = "plot")+geom_crossbar(data=d.plot.sum, aes(ymin=days,ymax=days),size=0.8,width=0.4)

library(gridExtra)
grid.arrange(p1t, p2t, nrow=2)

```

```{r additional-checks}
table(new.all.capt$sex.cat,new.all.capt$no.recaptures)

```


### New analyses of males' movement

 Because of critics, lets see what if we use body size and tusk length as a continuous variable in our models: 


```{r}
str(all.capt)
str(all.dist)
```

```{r}
str(new.all.capt.m)

new.all.capt.m$proc.rat<-new.all.capt.m$Tusk/new.all.capt.m$B.width

new.all.dist.m<-new.all.dist[new.all.dist$sex=="m",]

new.all.dist.m<-new.all.dist[new.all.dist$proc>0,]

new.all.dist.m
```

#### New analysis of number of recaptures

 Lets see the models:

```{r}


new.male.rec.nb<-glm.nb(no.recaptures~proc.rat,new.all.capt.m)
summary(new.male.rec.nb)

drop1(new.male.rec.nb,test="Chisq")
```


```{r}
plot(proc.rat~no.recaptures,new.all.capt.m)

```

```{r}
male.capt<-emtrends(new.male.rec.nb,~1,"proc.rat")
test(male.capt)
```

 Lets see the effect size
```{r}
rsq.partial(new.male.rec.nb,type='kl')
```

#### New analysis of number of days between the first and last encounter

```{r}
new.all.capt.m2<-new.all.capt.m[new.all.capt.m$no.recaptures>0,]

new.male.day.nb<-glm.nb(days~proc.rat,new.all.capt.m2)
summary(new.male.day.nb)

drop1(new.male.day.nb,test="Chisq")
```

```{r}
plot(proc.rat~days,new.all.capt.m2)

```

#### New analysis of total travelled distance

```{r}

new.male.dist<-lm(log(tot.dist+1)~proc.rat*no.recaptures,new.all.capt.m2)
drop1(new.male.dist,test="F")
```
```{r}
new.male.dist2<-update(new.male.dist,.~.-proc.rat:no.recaptures)
summary(new.male.dist2)
drop1(new.male.dist2,test="F")
```


```{r}
tot.dist.ph<-emtrends(new.male.dist2,~1,"proc.rat")
test(tot.dist.ph)
plot(tot.dist.ph)

tot.dist.ph2<-emmeans(new.male.dist2,~no.recaptures)
test(tot.dist.ph2)


```

#### New analysis of daily travelled distance

```{r}
summary(new.all.capt.m2$tot.dist)
summary(new.all.capt.m2$dist.a.day)

new.male.dist.d<-lm(log(tot.dist+1)/(days+1)~proc.rat*no.recaptures,
                    new.all.capt.m2)
drop1(new.male.dist,test="F")
```


```{r}
new.male.dist.d2<-update(new.male.dist.d,.~.-proc.rat:no.recaptures)
drop1(new.male.dist.d2,test="F")
```


 Lets build the mixed model and use distances between two encounters as a response variable

```{r}
new.all.dist.m$proc.rat<-new.all.dist.m$proc/new.all.dist.m$body

library(glmmTMB)
dist.new.m.tmb<-glmmTMB(log(Distance+1) ~ (proc.rat*day)^2+diff.days+(1|id),family="gaussian",data=new.all.dist.m)
summary(dist.new.m.tmb)
drop1(dist.new.m.tmb,test="Chisq")
```
```{r}
dist.new.m.tmb2<-update(dist.new.m.tmb,.~.-proc.rat:day)
summary(dist.new.m.tmb2)
drop1(dist.new.m.tmb2,test="Chisq")
```


```{r}
dist.ph<-emtrends(dist.new.m.tmb2,~1,"proc.rat")
test(dist.ph)

dist.ph2<-emtrends(dist.new.m.tmb2,~1,"diff.days")
test(dist.ph2)


plot(proc.rat~log(Distance+1),new.all.dist.m)
```

Using male body and tusk size ratio as a continous variable does not look well. We have significant results, but information loss as well. 

### Investigating above-ground activities in males at the moment of capture

We standardized the activity comments/variable the data in excel.

```{r}
tabi<-read.csv(file="/media/hanna/HDD/Projektek/Lethrus/susa.cap.recap.2016/activ.male.cln.csv",sep=",")

head(tabi)
str(tabi)
table(tabi$body.size)


```

 As the body size categories were added a long time ago, before the body size breakpoints were determined, we will use the sex.cat variable instead of body.size.

```{r}

tabi$

tabi$id<-as.character(tabi$id)

table(tabi$body.size,tabi$activity)
nrow(tabi)

table(tabi$megjegy)

tabi2<-tabi[tabi$megjegy!="NEWN",]

nrow(tabi2)

table(tabi2$body.size,tabi2$activity) #body size here is determined by sight and estimation not by test
table(tabi2$body.size)

library(dplyr)
library(ggplot2)


tabi3<-table(tabi2$body.size,tabi2$activity)

```

 Lets compare the individual IDs with the capture-recapture data. There should be a difference as we excluded individuals without coordinates during the analyses.

```{r}

i <- tabi2$id %in% new.all.capt.m$ID
ii <- new.all.capt.m$ID %in% tabi2$id
(missing.from.nalc <- as.character(tabi2$id[!i]))

(missing.from.tabi <- new.all.capt.m$ID[!ii])
```

```{r}

head(new.all.capt.m)
nrow(new.all.capt.m)

summary(tabi2$id)

```


```{r}
nrow(tabi2)
nrow(new.all.capt.m)

tabi222<-tabi2[,c(1,3,4,8,9,10)]
head(tabi222)
names(tabi222)[1]<-"ID"

nalc<-new.all.capt.m[,c(1,2,12)]

behav.m<-merge(nalc,tabi222,by=c("ID","sex"),all.x=TRUE,all.y=FALSE)
behav.m

nrow(behav.m)
nrow(nalc)

```


```{r}
(bh.t<-table(behav.m$activity,behav.m$sex.cat)) # sex.cat shows the body size categories based on the 2 breakpoint segmented model
```


```{r}

tabi.db <-as.data.frame(table(behav.m$sex.cat,behav.m$activity))

summary(tabi.db)

names(tabi.db)[1] <-"sex.cat"
names(tabi.db)[2] <-"act"
names(tabi.db)[3] <-"counts"

ggplot(tabi.db, aes(fill=act, y=counts, x=sex.cat)) + 
    geom_bar(position="fill", stat="identity")


chisq<-chisq.test(bh.t)

chisq

round(chisq$residuals,3)

library(corrplot)

corrplot(chisq$residuals,is.cor = FALSE)



chisq2<-chisq.test(bh.t, simulate.p.value = TRUE)
round(chisq$expected,2)

chisq2

chisq


round(chisq$expected,2)

```

 2024. 02. 28. We should compare the proportions of activities recorded within the 3 male categories.:
 We have no activity data from 2 individuals.
 Lets compare the proportions of the categories:

```{r}
bh.tn<-prop.table(bh.t,margin=1)
bh.tn

chisq.test(bh.tn)

```



There is an association between the season and the male morphs appearance which is not surprising as in march in individuals emerges and in april the reproductive season starts.


#### Using multinomial models for investigating above-ground activities in male morphs

```{r}
library(nnet)

behav.m

behav.m$date<-as.Date(behav.m$date)

behavm.mar<-behav.m %>% filter (behav.m$date<"2016-04-01")
behavm.mar$season<-"mar"

behavm.apr<- behav.m %>% filter (between(behav.m$date, as.Date("2016-04-01"), as.Date("2016-05-01")))
behavm.apr$season<-"apr"

behavm.maj<-behav.m %>% filter (behav.m$date>"2016-04-30")
behavm.maj$season<-"maj"

behav.mm<-rbind(behavm.mar,behavm.apr,behavm.maj)

behav.mm$date2

behav.mc<-aggregate(sex.cat~factor(sex.cat)+season+activity,behav.mm,FUN=length)
str(behav.mc)

behav.mc$season<-as.factor(behav.mc$season)
behav.mc$activity<-as.factor(behav.mc$activity)

names(behav.mc)[names(behav.mc) == "factor(sex.cat)"] <- "male.cat"
names(behav.mc)[names(behav.mc) == "sex.cat"] <- "Freq"

colnames(behav.mc)

with(behav.mc,table(male.cat,activity))
```

UPDATE: For the further analyses, lets use the other database the behav.m one as we do not want to count with frequencies.


Lets choose small sized males as a baseline. Furthermore, the multinom package does not include p-value calculation for the regression coefficients, so we calculate p-values using Wald tests (here z-tests).

```{r}
summary(behav.mm)
behav.mm$sex.cat<-as.factor(behav.mm$sex.cat)
behav.mm$activity<-as.factor(behav.mm$activity)
behav.mm$season<-as.factor(behav.mm$season)
behav.mm$sex.cat2<-relevel(behav.mm$sex.cat, ref="small.male")
behav.mm$season2<-relevel(behav.mm$season, ref="mar")
behav.mm$activity2<-relevel(behav.mm$activity,ref="mating")


```


Now the models:
```{r}
library(nnet)
test.multi<-multinom(sex.cat2~season2+activity2,data=behav.mm)
summary(test.multi)


test.multi2<-multinom(sex.cat2~(season2*activity2)^2,family=poisson,data=behav.mm)
summary(test.multi2)

anova(test.multi,test.multi2)
AIC(test.multi,test.multi2)

```

It seems, test.multi, the additive model fits better, but the difference is not significant. Lets continue with the model containing interaction.

Lets do the 2 tailed z test
```{r}
z <- summary(test.multi2)$coefficients/summary(test.multi2)$standard.errors
z

p <- (1 - pnorm(abs(z), 0, 1)) * 2
p

```




Just like binary logistic regression, we need to convert the coefficients to odds by taking the exponential of the coefficients.

```{r}
exp(coef(test.multi2))
```

```{r}
head(round(fitted(test.multi2), 2))
```

Analyses above give us not too much information, as the baseline used by the model is equal to observed frequency of mating in small males in March. One possible solution to do the tests separated for each month. Another idea to transform the Date into a numeric variable (Julian date). Lets do the last option.

```{r}
behav.mm$j.date<-as.numeric(format(behav.mm$date, format = "%j"))-76 # number of days since the start of the experiment

test.multij<-multinom(sex.cat2~j.date+activity2,dat=behav.mm)
summary(test.multij)


test.multij2<-multinom(sex.cat2~(j.date+activity2)^2,data=behav.mm)
summary(test.multij2)

anova(test.multij,test.multij2)
AIC(test.multij,test.multij2)

```

```{r}

z2 <- summary(test.multij2)$coefficients/summary(test.multij2)$standard.errors
z2


p2 <- (1 - pnorm(abs(z2), 0, 1)) * 2
p2

```
```{r}
studyl<-c(1:68)
summary(studyl)
```


We conducted the experiment for 68 days. According to that the first qu. is 18, the second qu. is 35 and the third qu. is 51. 
In the previous version quartiles were given in the emmeans using "at=list(j.date=(1:3*0.25))", but this setting put the quartiles to day16,day26,day41 because of the missing survey days. 
Taking the 18th, 35th and 51th days we get information about the 25%, 50%, and 75% days ellapsed since the start of the season/experiment.
```{r}

phmulti<-emmeans(test.multij2, ~sex.cat2 |j.date * activity2, at=list(j.date=c(18,35,51)))
test(phmulti)
tab.ph<-contrast(phmulti, method="pairwise",adjust="tukey")
pairs(phmulti)

summary(tab.ph)

comparison_table <- as.data.frame(summary(tab.ph))
write.csv(tab.ph, "emmeans_quantile_comparisons.csv", row.names = FALSE)



```


Lets make a plot showing activity frequences per male morph:

```{r}

library(grid)
library(gridExtra)


sm<-behav.mm[behav.mm$sex.cat=="small.male",]

nrow(sm)

sm.db<-data.frame(with(sm, table(j.date, activity)))

sm.db$perc<-(sm.db$Freq/111)*100


smed<-behav.mm[behav.mm$sex.cat=="middle.male",]

smed.db<-data.frame(with(smed, table(j.date, activity)))

smed.db$perc<-(smed.db$Freq/328)*100


smed.dat<-smed.db[,1,drop=FALSE]
smed.datt<-distinct(smed.dat)
smed.datt

sm.m<-data.frame(j.date=c("1","2","7","8","17","22","30","38","41","43","51","62","68"),
                          activity="dig",Freq=0,perc=0)

sm.dbb<-rbind(sm.db,sm.m)
as.data.frame(sm.dbb)

sorted_labels <- paste(sort(as.integer(levels(sm.dbb$j.date))))

sm.dbb$j.date <- factor(sm.dbb$j.date, levels = sorted_labels)

sm.dbbb<-with(sm.dbb,sm.dbb[order(sm.dbb$j.date),])




cols2<- c(dig="#d17341",feed_on="#918257",fight="#4f8286",mating="#912b1b", 
          tunnel="#f6e7b9", vegetation="#425f47")

require(ggplot2)

smp1<-ggplot(sm.dbbb, aes(x=j.date, y=perc, group = activity, color=activity,fill=activity))+ 
  geom_col(colour="black",width= 0.5) + 
  labs(title="c) Small males",x=" ",y=" ")+
  theme_classic()+theme(plot.title = element_text(size=13, face="bold"),legend.position="none") + scale_y_continuous(limits=c(0,8.5),breaks=seq(0,8.5,1))+ scale_fill_manual(values=cols2,labels=c('Feeding',"Fighting","Attempting to mate", "Being the entrance of a tunnel","Being in vegetation"))

smp1

smp11<-smp1+ labs(x="Number of days since start of the experiment", y="")+ theme(text=element_text(size=20))+theme (axis.text.x = element_text(face = "bold", size=12),  axis.text.y = element_text(face = "bold", size = 12),panel.border = element_rect(colour = "black", fill=NA, size=0.2))


smedp1<-ggplot(smed.db, aes(x=j.date, y=perc, group = activity,color=activity,fill=activity))+ 
  geom_col(colour="black",width= 0.5)+
  labs(title="b) Medium males",x=" ", y = "Percent of individuals (%)")+ theme_classic()+theme(plot.title = element_text(size=13, face="bold"),legend.position="none")+ scale_y_continuous(limits=c(0,8.5),breaks=seq(0,8.5,1)) + scale_fill_manual(values=cols2,labels=c('Feeding',"Fighting","Attempting to mate", "Being the entrance of a tunnel","Being in vegetation"))

smedp1

smedp11<-smedp1 + labs(y="Percent of individuals (%)", x="") + theme(text=element_text(size=20))+theme (axis.text.x = element_text(face = "bold", size=12),  axis.text.y = element_text(face = "bold", size = 12),panel.border = element_rect(colour = "black", fill=NA, size=0.2))
smedp11

sl<-behav.mm[behav.mm$sex.cat=="large.male",]

sl.db<-data.frame(with(sl, table(j.date, activity)))

sl.db$perc<-(sl.db$Freq/395)*100



sl.m<-data.frame(j.date=c("38"),
                 activity="dig",Freq=0,perc=0)

sl.dbb<-rbind(sl.db,sl.m)
as.data.frame(sl.dbb)

sorted_labels <- paste(sort(as.integer(levels(sl.dbb$j.date))))

sl.dbb$j.date <- factor(sl.dbb$j.date, levels = sorted_labels)

sl.dbbb<-with(sl.dbb,sl.dbb[order(sl.dbb$j.date),])



slp1<-ggplot(sl.dbbb, aes(x=j.date, y=perc, group = activity, color=activity,fill=activity))+ 
  geom_col(colour="black",width= 0.5)+
  labs(title="a) Large males",x="",y="")+
  theme_classic()+
  theme(plot.title = element_text(size=13, face="bold"))+ 
  scale_y_continuous(limits=c(0,8.5),breaks=seq(0,8.5,1))+ scale_fill_manual(values=cols2,labels=c("Digging", "Feeding","Fighting","Attempting to mate", "Being at the entrance","Being in vegetation"))

slp11<-slp1+labs(x="", y="") + guides(fill=guide_legend(title="Observed behaviour"))+ theme(text=element_text(size=20))+theme (axis.text.x = element_text(face = "bold", size=12),  axis.text.y = element_text(face = "bold", size = 12),panel.border = element_rect(colour = "black", fill=NA, size=0.2))

legend<-get_legend(slp11)


slp11 <- slp11 + theme(legend.position="none")


blank <- grid.rect(gp=gpar(col="white"))

```

```{r, fig.width=12, fig.height = 9}

fig3a<-grid.arrange(slp11,blank,smedp11,legend,smp11,blank,nrow=3,ncol=2, widths=c(2.5, 0.8),heights = c(3, 3,3))

ggsave("seasonal.movement.Figure.3a.tiff", fig3a, units="px",height= 4500, width= 4000,dpi=300)

```


```{r end}
sessionInfo()

```