---
title: "Analyses for Susa (2016) mark-recapture study"
author: "Johanna Kiss et al."
date: "2023-12-07"
output:
  html_document:
    toc: yes
    toc_depth: 4
    theme: united
  pdf_document: 
    toc: yes
    toc_depth: 4
    fig_caption: yes
    keep_tex: yes
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 18px;
  color: Gunmetal;
}
h1.title {
  font-size: 32px;
  color: Gunmetal;
  font-weight: bold;
}
h1 { /* Header 1 */
  font-size: 30px;
  color: BlueMossGreen;
}
h2 { /* Header 2 */
    font-size: 26px;
  color: DarkSlateGray;
  font-weight: bold;
}
h3 { /* Header 3 */
  font-size: 20px;
  color: DeepTeal;
  font-weight: bold;
}
h4 {/* Header 4 */
  font-size:16px;
  color: MediumTeal;
  font-weight: bold;
}
code.r{ /* Code block */
    font-size: 10px;
    color: CadetBlue;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 10px;
}
</style>


### Describing movement of different sized males and females in Lethrus beetles on field by using one model.

##### Using 1 model for each response variable (e.g., number of recaptures, travelled distances...) instead of 2.

##### To do that we reanalyse data and test the effect of the created 4-level factor ( female, small male, middle male, large male)


```{r}
all.dist<-read.csv(file="rel.full.data.csv",header=TRUE,nrow=FALSE)
all.capt<-read.csv(file="rel.aa.df.data.csv",header=TRUE,nrow=FALSE)
```

```{r}
head(all.dist)
head(all.capt)
```


##### Previously,  based on a segmented regression on body size we determined the breakpoints, now we use them to create male categories:

```{r}
library(dplyr)
new.all.capt<-all.capt%>%
  mutate(sex.cat= case_when(sex=="m" & B.width <= 10.68 ~ 'small.male',
                            sex=="m" & B.width > 12.32 ~ 'large.male',
                            sex=="m" & B.width > 10.68 & B.width <= 12.32 ~ 'middle.male',
                            sex=="f" ~ 'female'))

```

```{r}

library(dplyr)
new.all.dist<-all.dist%>%
  mutate(sex.cat= case_when(sex=="m" & body <= 10.68 ~ 'small.male',
                            sex=="m" & body > 12.32 ~ 'large.male',
                            sex=="m" & body > 10.68 & body <= 12.32 ~ 'middle.male',
                            sex=="f" ~ 'female'))

```

##### Load some packages:

```{r}
library(plotrix)
library(emmeans)
library(plyr)
library(dplyr)
library(ggplot2)
library(scales)
library(MASS)
library(lme4)
library(pbkrtest)
library(pscl)
library(lattice)
library(partR2)
library(rsq)
library(segmented)
library(ggplot2)
library(interactions)
library(jtools)
library(bbmle)
```


#### New analyses on the number of recaptures

```{r}
str(new.all.capt)
```

```{r}
new.all.capt.m<-new.all.capt[new.all.capt$sex=="m",]

new.all.capt.m<-new.all.capt[!new.all.capt$Tusk==0,]




custom_pch <- c(3,15,17)

plot(new.all.capt.m$Tusk~new.all.capt.m$B.width, type="n", xlab="Body width [mm]", ylab="Tusk length [mm]",las=1,  cex.axis=1.3, cex=1, cex.lab=1.7, las=1, font.lab=2.3, font.main=4.3, ylim=c(0, 5),xlim=c(9,14))
points(new.all.capt.m$Tusk~new.all.capt.m$B.width, subset=new.all.capt.m$sex.cat=="small.male", pch=3,
       col=1)
points(new.all.capt.m$Tusk~new.all.capt.m$B.width, subset=new.all.capt.m$sex.cat=="middle.male", pch=15,
       col="grey")

points(new.all.capt.m$Tusk~new.all.capt.m$B.width, subset=new.all.capt.m$sex.cat=="large.male", pch=17,
       col="darkgrey")

abline(v=c(10.68, 12.32), col=c("darkgrey", "black"), lty=2, lwd=2)


```

##### Lets see the models:

```{r}
new.p.rec<-glm(no.recaptures~sex.cat,new.all.capt,family=poisson)
drop1(new.p.rec,test="Chisq")
```

```{r}
new.qp.rec<-glm(no.recaptures~sex.cat,new.all.capt,family=quasipoisson)
drop1(new.qp.rec,test="Chisq")
```

```{r}
new.zp.rec<-zeroinfl(no.recaptures~sex.cat,new.all.capt,dist='poisson')
drop1(new.zp.rec,test="Chisq")
```

```{r}
new.znb.rec<-zeroinfl(no.recaptures~sex.cat,new.all.capt,dist='negbin')
drop1(new.znb.rec,test="Chisq")
```
```{r}
new.hnb.rec<-hurdle(no.recaptures~sex.cat,new.all.capt,dist='negbin')
drop1(new.hnb.rec,test="Chisq")
```

```{r}
new.nb.rec<-glm.nb(no.recaptures~sex.cat,new.all.capt)
drop1(new.nb.rec,test="Chisq")
```

##### Lets compare the models

```{r}

logLik(new.p.rec)
logLik(new.nb.rec)
logLik(new.hnb.rec)
logLik(new.znb.rec)
logLik(new.zp.rec)
logLik(new.qp.rec)
```
```{r}

AIC(new.p.rec)
AIC(new.nb.rec)
AIC(new.hnb.rec)
AIC(new.znb.rec)
AIC(new.zp.rec)
AIC(new.qp.rec)
```
##### HNB and NB models seemed to be the best.

```{r}
drop1(new.hnb.rec,test="Chisq")
drop1(new.nb.rec,test="Chisq")
```

```{r}
ph.test.rec<-emmeans(new.nb.rec, ~sex.cat,type="response")
test(ph.test.rec)
pairs(ph.test.rec)

plot(ph.test.rec)


ph.test.rec2<-emmeans(new.nb.rec, ~sex.cat,type="link")
test(ph.test.rec2)
pairs(ph.test.rec2)

plot(ph.test.rec)



```

```{r}

boxplot(new.all.capt$no.recaptures~new.all.capt$sex.cat,xlab="Sex",ylab="Number of (re)captures",las=1, lwd=1,font.lab=2, cex.lab=1.3,cex.axis=1.3,font.main=2)

new1<-ggplot(new.all.capt,aes(sex.cat,no.recaptures),line.thickness=1.2) +theme_classic(base_size = 18) + geom_jitter(width=0.2,alpha=0.5)+scale_y_continuous(name="Number of (re)captures",breaks= pretty_breaks()) +theme (axis.text.x = element_text(face = "bold", size=12,colour="black"), axis.text.y = element_text(face = "bold", size = 12,colour="black"),panel.border= element_rect(color = "black",fill=NA,size=1), legend.text=element_text(size=14),legend.title=element_text(size=16))

```

```{r fig.height = 10, fig.width = 10, fig.align = "center"}

new.plot1<-new.all.capt[,c(8,13)]
new.plot1.sum<-aggregate(.~sex.cat,mean,data=new.plot1)



new1+ theme(plot.title.position = "plot")+geom_crossbar(data=new.plot1.sum, aes(ymin=no.recaptures,ymax=no.recaptures),size=0.8,width=0.4)

```

##### Lets see now the effect size:

```{r}
rsq.partial(new.nb.rec,type='kl')
```

##### Females and male morphs showed a small effect size: 0.014.

##### We can calculate R2 values for Hurdle models by using the function r2_zeroinflated of performance package. It is used for models with zero-inflation component, including mixed effects models. Hurdle and zeroinfl functions by Achim Zeileis

```{r}
library(performance)
r2_zeroinflated(new.hnb.rec)
```


#### New analyses on the number of days


##### Lets see the models:

```{r}
new.p.days<-glm(days~(sex.cat*no.recaptures)^2,new.all.capt,family=poisson,subset=new.all.capt$no.recaptures>0)
drop1(new.p.days,test="Chisq")

new.p.days1<-update(new.p.days,.~.-sex.cat:no.recaptures)
drop1(new.p.days1,test="Chisq")
```

```{r}
new.qp.days<-glm(days~(sex.cat*no.recaptures)^2,new.all.capt,family=quasipoisson,subset=new.all.capt$no.recaptures>0)
drop1(new.qp.days,test="Chisq")

new.qp.days1<-update(new.qp.days,.~.-sex.cat:no.recaptures)
drop1(new.qp.days1,test="Chisq")
```

```{r}
new.zp.days<-zeroinfl(days~(sex.cat*no.recaptures)^2,new.all.capt,dist='poisson',subset=new.all.capt$no.recaptures>0)
drop1(new.zp.days,test="Chisq")

new.zp.days1<-update(new.zp.days,.~.-sex.cat:no.recaptures)
drop1(new.zp.days1,test="Chisq")

```

```{r}
new.znb.days<-zeroinfl(days~(sex.cat*no.recaptures)^2,new.all.capt,dist='negbin',subset=new.all.capt$no.recaptures>0)
drop1(new.znb.days,test="Chisq")

new.znb.days1<-update(new.znb.days,.~.-sex.cat:no.recaptures)
drop1(new.znb.days1,test="Chisq")
```
```{r}
new.hnb.days<-hurdle(days~(sex.cat*no.recaptures)^2,new.all.capt,dist='negbin',subset=new.all.capt$no.recaptures>0)
drop1(new.hnb.days,test="Chisq")

new.hnb.days1<-update(new.hnb.days,.~.-sex.cat:no.recaptures)
drop1(new.hnb.days1,test="Chisq")
```

```{r}
new.nb.days<-glm.nb(days~(sex.cat*no.recaptures)^2,new.all.capt,subset=new.all.capt$no.recaptures>0)
drop1(new.nb.days,test="Chisq")

new.nb.days1<-update(new.nb.days,.~.-sex.cat:no.recaptures)
drop1(new.nb.days1,test="Chisq")
```

##### Lets compare the models

```{r}

logLik(new.p.days1)
logLik(new.nb.days1)
logLik(new.hnb.days1)
logLik(new.znb.days1)
logLik(new.zp.days1)
logLik(new.qp.days1)
```

```{r}

AIC(new.p.days1)
AIC(new.nb.days1)
AIC(new.hnb.days1)
AIC(new.znb.days1)
AIC(new.zp.days1)
AIC(new.qp.days1)
```
##### Here also HNB and NB models seemed to be the best.

```{r}
drop1(new.hnb.days1,test="Chisq")
drop1(new.nb.days1,test="Chisq")
```
##### Its seems that sex and male morphs had no significant effect on the number of days individuals were seen on field.


```{r}

new.all.capt2<-new.all.capt[new.all.capt$no.recaptures>0,]


boxplot(new.all.capt2$days~new.all.capt2$sex.cat,xlab="Sex",ylab="Number of days between the first and last encounter",las=1, lwd=1,font.lab=2, cex.lab=1.3,cex.axis=1.3,font.main=2)
```


```{r}

new2<-ggplot(new.all.capt2,aes(sex.cat,days),line.thickness=1.2) +theme_classic(base_size = 18) + geom_jitter(width=0.2,alpha=0.5)+scale_y_continuous(name="Number of days between the first and last encounter",breaks= pretty_breaks()) +theme (axis.text.x = element_text(face = "bold", size=12,colour="black"), axis.text.y = element_text(face = "bold", size = 12,colour="black"),panel.border= element_rect(color = "black",fill=NA,size=1), legend.text=element_text(size=14),legend.title=element_text(size=16))

```

```{r fig.height = 8, fig.width = 8, fig.align = "center"}


new.plot2<-new.all.capt2[,c(6,13)]
new.plot2.sum<-aggregate(.~sex.cat,mean,data=new.plot2)



new2+ theme(plot.title.position = "plot")+geom_crossbar(data=new.plot2.sum, aes(ymin=days,ymax=days),size=0.8,width=0.4)

```

##### Lets see now the effect size:

```{r}
rsq.partial(new.nb.days1,type='kl')
```


##### We can calculate R2 values for Hurdle models by using the function r2_zeroinflated of performance package. It is used for models with zero-inflation component, including mixed effects models. Hurdle and zeroinfl functions by Achim Zeileis

```{r}
library(performance)
r2_zeroinflated(new.hnb.days1)
```

```{r}
r2(new.hnb.rec)
```

##### Females and male morphs showed also a small effect size: 0.018.

```{r}
ph.test.days<-emmeans(new.nb.days1, ~sex.cat,type="response")
test(ph.test.days)
pairs(ph.test.days)
contrast(ph.test.days)

plot(ph.test.days)


ph.test.days2<-emmeans(new.nb.days1, ~sex.cat,type="link")
test(ph.test.days2)
pairs(ph.test.days2)
contrast(ph.test.days2)

plot(ph.test.days2)
```


#### New analyses of distances travelled between two encounters


##### Response variables: female-male categories (4-level factor), number of days since start of the study, number of days between two encounters, and id as a random factor.

```{r}
colnames(new.all.dist)
str(new.all.dist)
```
```{r}
library(lme4)
dist.new<-lmer(log(Distance+1) ~ (sex.cat+day+diff.days)^2+(1|id),data=new.all.dist)
print(summary(dist.new),correlation=TRUE)
drop1(dist.new,test="Chisq")
```






```{r}
dist.new2<-update(dist.new,.~.-sex.cat:diff.days)
summary(dist.new2)
drop1(dist.new2,test="Chisq")
```


```{r}
dist.new3<-update(dist.new2,.~.-day:diff.days)
summary(dist.new3)
drop1(dist.new3,test="Chisq")
```
Compare "stats drop1" to "glmmTMB drop1":

```{r}
library(glmmTMB)
dist.new.tmb<-glmmTMB(log(Distance+1) ~ sex.cat:day+diff.days+(1|id),family="gaussian",data=new.all.dist)
summary(dist.new.tmb)
drop1(dist.new.tmb,test="Chisq")
```

```{r}
ph.dist.new<-emmeans::emtrends(dist.new3,~1,"diff.days")
test(ph.dist.new)
```

```{r}
effect_plot(dist.new3, pred = diff.days, interval = FALSE, plot.points = TRUE,x.label="Number of days between encounters")+ theme_apa()
```

```{r}
ph.dist.new2<-emtrends(dist.new3,~sex.cat,"day")
test(ph.dist.new2)

pairs(ph.dist.new2)

emmip(dist.new3,sex.cat~day,cov.reduce=range)
```
##### Small males seem to travel and move similarly to the females.

##### Lets visualise the results.

```{r fig.height = 8, fig.width = 14, fig.align = "center"}


new.dist1<-interact_plot(dist.new3, pred = day, modx = sex.cat,x.label="Number of days since the start of the study",data=new.all.dist,y.label="Log(Distance+1)",lwd=2,line.thickness=1.2,colors=c("black","grey35","grey55","grey6"),plot.points=TRUE,point.shape=T,point.size=3,interval = TRUE,  int.type = "confidence", int.width = .95,modx.labels=c("Female","Large male","Medium male","Small male")) + theme_apa()+theme_classic(base_size = 18)+theme(legend.title=element_blank(),legend.key.size = unit(3, "lines"),legend.position="right",legend.justification = "top")+guides(fill = guide_legend(override.aes = list(size=4)))+ scale_x_continuous(limits=c(0.04,71),expand=c(0,0),breaks=c(10,20,30,40,50,60)) + scale_y_continuous(limits=c(0.04,6.4), expand=c(0,0),breaks=c(1,2,3,4,5,6))



new.dist1 + theme (axis.text.x = element_text(face = "bold", size=12),  axis.text.y = element_text(face = "bold", size = 12),panel.border = element_rect(colour = "black", fill=NA, size=0.2))
```

##### Create colourfull plots for graphical abstract:

```{r fig.height = 8, fig.width = 14, fig.align = "center"}


new.dist.col<-interact_plot(dist.new3, pred = day, modx = sex.cat,x.label="Number of days since the start of the study",data=new.all.dist,y.label="Log(Distance+1)",lwd=2,line.thickness=1.2,colors=c("green","brown","red","orange"),interval = TRUE, int.type = "confidence", int.width = .95) + theme_apa()+theme_classic(base_size = 18)+ scale_x_continuous(limits=c(0.04,71),expand=c(0,0),breaks=c(10,20,30,40,50,60)) + scale_y_continuous(limits=c(0.04,6.4), expand=c(0,0),breaks=c(1,2,3,4,5,6))+theme(legend.position = "none")



new.dist.col + theme (axis.text.x = element_text(face = "bold", size=16),  axis.text.y = element_text(face = "bold", size = 16),panel.border = element_rect(colour = "brown", fill=NA, size=0.5))




new.dist.col2<-interact_plot(dist.new3, pred = day, modx = sex.cat,x.label="Number of days since the start of the study",data=new.all.dist,y.label="Log(Distance+1)",lwd=2,line.thickness=1.2,colors=c("darkolivegreen4","brown4","coral2","darkgoldenrod2"),,plot.points=TRUE,point.shape=T,point.size=3,interval = TRUE,  int.type = "confidence", int.width = .95,modx.labels=c("Female","Large male","Medium male","Small male")) + theme_apa()+theme_classic(base_size = 18)+theme(legend.title=element_blank(),legend.key.size = unit(3, "lines"),legend.position="right",legend.justification = "top")+guides(fill = guide_legend(override.aes = list(size=4)))+ scale_x_continuous(limits=c(0.04,71),expand=c(0,0),breaks=c(10,20,30,40,50,60)) + scale_y_continuous(limits=c(0.04,6.4), expand=c(0,0),breaks=c(1,2,3,4,5,6))



new.dist.col2 + theme (axis.text.x = element_text(face = "bold", size=12),  axis.text.y = element_text(face = "bold", size = 12),panel.border = element_rect(colour = "black", fill=NA, size=0.2))


```


##### Lets see effect sizes:

```{r}
partR2(dist.new3, partvars= c("diff.days","sex.cat:day"),R2_type="conditional",nboot = 100, CI = 0.95)

```

```{r}
partR2(dist.new3, partvars= c("diff.days","sex.cat:day"),R2_type="marginal",nboot = 100, CI = 0.95)
```

##### Semi-partial R2 of sex-size category and number of days since start of the study: marginalR2= 0.035, conditionalR2=0.155, i.e., representing medium-leveled effect.

##### Lets create a plot for the first two models:

```{r plot-fig-1-2}
d.plot1<-new.all.capt[,c(8,13)]
d.plot1.sum<-aggregate(.~sex.cat,mean,data=d.plot1)

p1new<-ggplot(d.plot1,aes(sex.cat,no.recaptures),line.thickness=1.2) +theme_classic(base_size = 18)+geom_point(position=position_jitter(width=0.2),size=3,aes(shape=sex.cat,x = sex.cat, y = no.recaptures,color=sex.cat))+
scale_y_continuous(name="Number of encounters",limits=c(0,11),breaks= round(seq(min(new.all.capt$no.recaptures),max(new.all.capt$no.recaptures),by=1),1)) +  labs(x="Sex")+scale_x_discrete(labels=c("Female","Large male","Medium male", "Small male"))+theme (axis.text.x = element_text(face = "bold", size=16,colour="black"), axis.text.y = element_text(face = "bold", size = 12),panel.border= element_rect(color ="black",fill=NA,size=1))+geom_crossbar(data=d.plot1.sum, aes(ymin=no.recaptures,ymax=no.recaptures),size=0.8,width=0.6)+theme(legend.position = "none")+ scale_color_manual(values = c("black","grey35","grey55","grey6"))

p1t<-p1new+labs(title = "a)") + theme(plot.title.position = "plot")+geom_crossbar(data=d.plot1.sum, aes(ymin=no.recaptures,ymax=no.recaptures),size=0.8,width=0.4)

d.plot<-new.all.capt2[,c(6,13)]
d.plot.sum<-aggregate(.~sex.cat,mean,data=d.plot)

p2<-ggplot(d.plot,aes(sex.cat,days),line.thickness=1.2) +theme_classic(base_size = 18)+geom_point(position=position_jitter(width=0.2),size=3,aes(shape=sex.cat,x = sex.cat, y = days, color = sex.cat))+
scale_y_continuous(name="Number of days",limits=c(0,57),breaks= round(seq(min(new.all.capt2$days),max(new.all.capt2$days),by=3),1)) +  labs(x="Sex")+scale_x_discrete(labels=c("Female","Large male","Medium male", "Small male"))+theme (axis.text.x = element_text(face = "bold", size=16,colour="black"), axis.text.y = element_text(face = "bold", size = 12),panel.border= element_rect(color ="black",fill=NA,size=1))+geom_crossbar(data=d.plot.sum, aes(ymin=days,ymax=days),size=0.8,width=0.6)+theme(legend.position = "none")+ scale_color_manual(values = c("black","grey35","grey55","grey6"))


p2t<-p2+labs(title = "b)") + theme(plot.title.position = "plot")+geom_crossbar(data=d.plot.sum, aes(ymin=days,ymax=days),size=0.8,width=0.4)

library(gridExtra)
grid.arrange(p1t, p2t, nrow=2)

```

```{r additional-checks}
table(new.all.capt$sex.cat,new.all.capt$no.recaptures)

```

```{r end}
sessionInfo()

```